---
title: "Par Inc. Golf Ball "
output: html_notebook
---

Problem Statement

Par Inc., is a major manufacturer of golf equipment. Management believes that Par’s market share could be
increased with the introduction of a cut-resistant, longer-lasting golf ball. Therefore, the research group at Par
has been investigating a new golf ball coating designed to resist cuts and provide a more durable ball. The
tests with the coating have been promising.

One of the researchers voiced concern about the effect of the new coating on driving distances. Par would like
the new cut resistant ball to offer driving distances comparable to those of the current-model golf ball. To
compare the driving distances for the two balls, 40 balls of both the new and current models were subjected to
distance tests.

The testing was performed with a mechanical hitting machine so that any difference between the mean
distances for the two models could be attributed to a difference in the design. The results of the tests, with
distances measured to the nearest yard, are contained in the data set “Golf”.

```{r}
library(readxl)
Par_Inc <- read_excel("Par-Inc.xls")
```


```{r}

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

