---
title: "Project - Cardio Good Fitness"
output: html_notebook
---
Project Objective

The objective of the report is to explore the cardio data set (“CardioGoodFitness”) in R
and generate insights about the data set. This exploration report will consists of the following:

Importing the dataset in R

Understanding the structure of dataset

Graphical exploration

Descriptive statistics

Insights from the dataset


Setting the working directory
```{r}
setwd("/home/rstudio/pgpbabi/Project1_Cardio_Data_Exploration")
```
Reading the csv data into r object
```{r}
cg_data <- read.csv('CardioGoodFitness.csv')
```
cg_data dimensions
```{r}
dim(cg_data)
```
Examining the first few rows
```{r}
head(cg_data)
```


Finding class of each feature along with internal structure

```{r}
str(cg_data)
```

The Cardio Good Fitness data has 3 features Product, Gender, Marital Status as factors.



Summary statistics
```{r}
summary(cg_data)
```
Inferences from the summary:

1. The dataset consists of data about 3 products

2. The dataset is for the agegroups of 18-50 years of bith male and female.

3. There are more male observations than female

4. Education seems to be a numeric value representing numberof years spent on education by cutomer
   All the customers seems to be having at least school level. Some of them have 21 years which is    roughly equivalent to doctorate 

5. Other features are self explanatory from the summary above


Data Exploration

1. Finding the fittest observations

```{r}
fittest <- cg_data[cg_data$Fitness==5,]
fittest
```

Out of 180 observations, 30 are the fittest. Roughly, 17% of total customers are fitness savvy

Let's find out the popular product among the fittest
```{r}
summary(fittest)
```

```{r}
str(fittest)
```

From the above results, 
1. Among the most fit customers, male tend to be more fitness savvy than females as there is a       clear difference of 25:6 (male:female) in numbers by gender.
2. The most popular product among ultra fit people is TM798

Examining the least fit set of observations
```{r}
least_fit<-cg_data[cg_data$Fitness==1,]
least_fit
```
Out of 180 observations, only 2 are least fit. There is no sufficient data about the least fit people as it's difficult to analyze with just 2 rows


Number of people by fitness levels
```{r}
table(cg_data$Fitness)
```

AMong all, almost half of them are moderately fit with fitness level 3.